Status of Global Unassessed Fisheries will Remain Highly Uncertain without Better Data

### 2021-01-16 documents/ovando-etal-assessing-global-fisheries-short.Rmd
getting to the point

The United Nations Sustainable Development Goal 14 (SDG 14), focusing on "Life under water", calls for the global community to "Conserve and sustainably use the oceans, seas and marine resources for sustainable development".While our understanding of the state of world fisheries has improved over the last decade the majority of the world's fish populations, making up roughly 50% of marine landings, lack formal statistical assessments (stock assessments) of their biological status [@hilborn2020;@fao2020]. Meeting the SDG 14 targets depends in part on improving our ability to effectively track the status of global marine fish populations.

The Food and Agriculture Organization of the United Nations' (FAO) is the custodian agency of the SDG 14 Indicator on fisheries sustainability, and the FAO's report on the State of World Fisheries and Aquaculture (SOFIA) report is the most widely used primary source for tracking the global state of fisheries. The latest SOFIA report, covering 70% of the landings of all fisheries in the world, estimates that as of 2017 59.6% of marine fish stocks are maximally sustainably fished, 6.2% are underfished, and 34.2% are overfished [@fao2020]. Where possible, statements about the status of individual fish stocks are made on the basis of formal stock assessments reported in the RAM Legacy Stock Assessment Database [@ricard2012]. The "assessed" fisheries in RAM represent roughly 50% of global capture as of 2020 [@hilborn2020], and represent our best estimates of the biological state of "assessed" fish populations around the globe.

That leaves roughly 50% of global landings, and the majority of global fisheries, as currently "unassessed". The SOFIA report includes unassessed stocks and bases its estimates for these fisheries on data-limited methods or qualitative expert opinion for each region where these stocks are distributed [@fao2020]. While these methods combined with local knowledge can be well informed as to the status of their fish stocks, the SOFIA assessment was designed in the 1970s based on the then available data and methods. With the surge in data availability, such as global assessments of management strength, trawl footprints, and fishing effort, and modeling methods,the SOFIA global assessment methods for unassessed fish populations need to be updated to meet the demand for tracking progress of global fish populations towards the SDG goals.

In this paper we evaluate the ability of alternative data-limited assessment models to improve our estimate of the population status of unassessed fish stocks around the world. While these unassessed stocks are generally individually smaller than the typically larger and more valuable fish populations in the assessed category, collectively they are a vital source of food, employment, cultural value, and ecosystem services around the world. Numerous studies in recent years have put forward versions of "data-limited" models that have attempted to provide estimates for the global status of unassessed fish stocks [@costello2012b; @costello2016a; @pauly2007; @rosenberg2018; @thorson2012a]. Due to data limitations, all of these global assessment efforts use forms of "catch-only" stock assessment models (Free et al. 2020 [@free2020] and references therein). These models seek to infer stock status, for example biomass B relative to the biomass at maximum sustainable yield B~MSY~, from characteristics of a fishery's catch history, for example the ratio of catch to maximum catch [@martell2013].

However, Free et al. 2020 [@free2020] demonstrated that these catch-only models can often produce both imprecise and biased estimates of current stock status in terms of B/B~MSY~. These issues become apparent when we consider some of the macro-level predictions made by these models. While the estimates B/B~MSY~ and other reference points reported in RAM are themselves model outputs subject to their own non-trivial errors and biases, a simple benchmark is to compare the estimates of fishery status from RAM to those predicted by potentially less reliable methods intended for use when insufficient data are available for a full stock assessment model.

Costello et al. 2016 [@costello2016a] finds similar rankings of regions in terms of stock status as RAM, but their estimate of state of fisheries in the Mediterranean/Black Sea regions and Southeast Asia seem to be over-optimistic, and the Northeast Pacific should be better by comparison. The superensemble method [@anderson2017a] used in Rosenberg et al. 2018 [@rosenberg2018] demonstrates the same problem, with stocks in Southeast Asia estimated as doing better than the Northeast Atlantic or Northeast Pacific. The Pauly 2007 [@pauly2007] catch-based approach finds the stocks of Southeast Asia in much better condition than those in the Northeast Pacific or Northeast Atlantic (Table.\@ref(tab:status-tab)). The non-RAM results in Table.\@ref(tab:status-tab) include both formally assessed and unassessed stocks, and as such we would expect them to differ broadly in their estimates of regional stock status, particularly since we might expect unassessed stocks to be less rigorously managed and by extension have poorer stock status. However, the lack of consistency across heavily assessed regions such as the Northeast Pacific, and the lack of contrast in stock status between heavily and lightly managed regions is concerning.

@@

readr::read_csv(here("data","status-table.csv")) %>% 
  select(-`B/BMSY RAM`) %>%  
  left_join(o_by_a, by = c("FAO Area" = "fao_area")) %>% 
  left_join(state_space_results, by = c("FAO Area" = "fao_area")) %>% 
  arrange(`FAO % Overfished`) %>% 
  select(`FAO Area`,`FAO % Overfished`, everything()) %>% 
  knitr::kable(caption = "Estimates of B/B~MSY~ by FAO, RAM Legacy Stock Assessment Database through Hilborn et al. 2020, Costello et al. 2016, 
               Rosenberg et al. 2018, and Pauly 2007",
               booktabs = TRUE,
               format = "pipe") %>% 
  kableExtra::kable_styling(full_width = T, font_size = 9) %>%
  kableExtra::row_spec(0, bold = T)

@@

In this paper we ask, can combining the FAO's catch statistics with other broadly available data improve our understanding of the state of global fish populations? We use a flexible surplus-production based stock assessment software package, [sraplus](www.github.com/danovando/sraplus) to demonstrate how different data sources can be used to augment catch-only models at a global scale, and to evaluate how our perception of global stock status would vary depending on which sources of data we include. We show that improving our understanding of the world's unassessed fisheries depends on a redoubled effort at global data collection and synthesis.

@@

We compared estimates of stock status, defined by B/B~MSY~ obtained from sraplus using various forms of broadly available data to the best-available estimates of B/B~MSY~ available for those fish populations reported in the RAM database. All analysis were conducted in the R programming language [@rcoreteam2019]. Model fitting was conducted using Rcpp [@eddelbuettel2011] and Stan [@carpenter2017] implemented through Template Model Builder [@kristensen2016] by the tmbstan package [@monnahan2018]. The sraplus package is publicly available at [github.com/danovando/sraplus](https://github.com/DanOvando/sraplus), and all materials needed to fully reproduce this manuscript are available at [github.com/DanOvando/assessing-global-fisheries](https://github.com/DanOvando/assessing-global-fisheries). We encourage readers to explore the documentation available at the package website at www.github.com/danovando/sraplus. Here we describe the structure of the population model underpinning sraplus, the estimation models used, and the construction of priors used in this paper.

## Data Sources

We restricted ourselves to databases with wide coverage around the world, representing the types of information that can be leveraged in efforts to assess the status of global fisheries (Table.\@ref(tab:data)). While fishery-independent surveys are becoming increasingly available [@maureaud2020], they are not yet sufficiently distributed or accessible to serve as a foundation for assessment of global unassessed fisheries.

@@

## Experiment Structure

We assessed the ability of Fisheries Management Index (FMI), swept area ratio (SAR), and effort data to improve estimates of the state of global fish populations. We based this test on `r ram_comp_data$stockid %>% n_distinct()` stocks from RAM, covering `r n_distinct(ram_comp_data$isscaap_group)` broad taxonomic groups, with estimates of B/B~MSY~ and greater than 25 years of continuous catch history. B/B~MSY~ values from RAM are themselves estimates, not data, but they are the best available information on global stock status. We then paired the catch histories for these RAM stocks with regional-level SAR, FMI, and effort data. This process approximates a global-level assessment exercise, where data are available at regional levels, but not for specific fisheries.

We also estimated B/B~MSY~ values of our candidate RAM stocks by using sraplus to fit to an abundance index drawn directly from RAM. We then fit a range of models utilizing different combinations FMI, SAR, and effort data, along with the CMSY method described in Froese et al. 2017 [@froese2017] (See Table.\@ref(tab:dat-desc)). We assessed performance using three metrics: median percent error (MPE, a measure of bias), median absolute percent error (MAPE, a measure of accuracy), and classification accuracy. Classification accuracy was calculated as the proportion of times that use of a given combination of data resulted in a stock being classified into the correct FAO status classification (one of underfished, maximally sustainably fished, and overfished).

@@

The core of sraplus is a Pella-Tomlinson [@pella1969] production model constructed in the manner of [@winker2018]. While models of these kinds abstract away many important details of fish biology and fleet behavior, they are the highest resolution model that the potential data evaluated here will support. The purpose of sraplus is not to make substantial innovations in the fitting of surplus production models, but to provide a flexible tool for exploring the impacts of adding different kinds of data and priors on estimates of fishery status

@@

We allow for process error *p* (in the manner of the stochastic stock reduction analysis error suggested by [@walters2006]). This allows the population dynamics to deviate from the exact values given by the Pella-Tomlinson operating model, while still conforming to the assumptions of this model on average. Incorporation of process errors is useful for two reasons: (1) when you have an abundance index, process errors can reduce bias arising from lack of fit in a deterministic SRA whenever dynamics are poorly explained by catch-history alone, and (2) with or without an abundance index (or other info), the stochastic portion is necessary to get good uncertainty intervals (i.e., with close to nominal coverage, see [@thorson2018]).

@@

In addition to the results from sraplus, we include a set of results produced by the CMSY method [@froese2017]. For computational efficiency, we used a ported version of the CMSY model available at <https://github.com/DanOvando/portedcmsy>. The only modification made is to convert the underlying population model to C++ for faster computation. For each stock we used all the default options provided by CMSY, except for resilience, which was pulled from the vulnerability scores from FishBase accessed through `rfishbase` [@boettiger2012]. Vulnerability scores greater than 66 were scored as low resilience, between 33 and 66 medium resilience, and lower than 33 high resilience.

In the absence of an index of abundance CMSY and sraplus are conceptually the same: Both sample from prior distributions of life history and stock status conditional on the stock not collapsing given a catch history and a functional form for the population growth equation. The core different between CMSY and sraplus as utilized here then is that when we refer to CMSY we mean CMSY run using the default heuristics used by the algorithm to construct priors on initial, intermediate, and terminal stock status. In contrast, sraplus fit using for example FMI data uses the empirical models described here to generate priors on stock status. In addition, all sraplus runs use a prior-predictive tuning procedure to resolve issues related to Borel's paradox, described below.

@@

Suppose that the only thing we observe from a fishery is a catch history. Assuming Pella-Tomlinson population dynamics, the only thing we can learn from this catch history alone is the set of model parameters that ensure that the population still exists and never collapsed in the past. We can think of this as a binomial process in which we fit a model of the population, conditional on catches, to the fact that we know that population existed in each time step of the catch history. Beyond that though, baring additional information or model assumptions we have no way of knowing whether these catches represent a substantial proportion of a small population or a minuscule fraction of a massive population; all we know is that the population must always have been above zero.

In the absence of any data to fit to, the SRA algorithm works by assuming that we know current stock status, and then finds feasible parameters to satisfy that belief. This creates a problem for the Bayesian nature of our analysis though. Consider a production model with two parameters, a growth rate *r* and a carrying capacity *K*. Once we specify prior distributions on *r* and *K*, and then apply these distributions to our model (the shape of the production function along with the catch histories), we have implicitly provided a prior on the status of the stock in all time periods, since each unique combination of *r* and *K* together with the model and the catch history produces a deterministic stock status in each time step. Doing so places essentially two priors on recent stock status: one implicit prior through the population parameter priors, and one explicit through the users perception of recent stock status, creating a problem termed Borel's Paradox (See Poole and Raftery 2000 [@poole2000] and references therein for a discussion of Borel's Paradox in a fisheries context). This may seem like an academic concern, and indeed in our experience when the data are sufficiently informative the Bayesian version of our model subject to Borel's paradox produces effectively identical results to those produce by the same model fit by maximum likelihood. However, particularly for the SRA version of sraplus, Borel's Paradox poses a particular problem.

@@

We first present a case study demonstrating how different kinds of data can lead to different conclusions about stock status. From there, we assess the performance of models fit using different kinds of broadly available data: combinations of catches, effort, Fisheries Management Index (FMI) scores (a measure of management capacity), and swept area ratio (SAR) values (a measure of trawl fishing intensity) (Table.\@ref(tab:data)).

>>>>

The United Nations Sustainable Development Goal 14 (SDG 14), focusing on "Life under water", calls for the global community to "Conserve and sustainably use the oceans, seas and marine resources for sustainable development". Marine fisheries represent one of the largest anthropenic impacts on the oceans, and when managed sustainably are a critical source of economic prosperity and food security around the globe. As such, meeting the SDG 14 targets depends in part on our ability to effectively measure the status of global marine fish populations and fisheries. While our understanding of the state of world fisheries has improved over the last decade the majority of the world's fish populations, making up roughly 50% of marine landings, lack formal statistical assessments (stock assessments) of their biological status [@hilborn2020;@fao2020]. This is a major impediment to ensuring the sustainable development of the world's oceans. In this paper, we assess why the assessment of global fisheries remains a challenge, and chart a path towards a better understanding of the state of the world's marine resources.

The Food and Agriculture Organization of the United Nations' (FAO) is the custodian agency of the SDG 14 Indicator on fisheries sustainability, and the FAO's State of World Fisheries and Aquaculture (SOFIA) report is the most widely used primary source for tracking the global state of fisheries. The latest SOFIA report, covering 70% of the landings of all fisheries in the world, estimates that as of 2017 59.6% of marine fish stocks are maximally sustainably fished, 6.2% are underfished, and 34.2% are overfished [@fao2020]. Where possible, these statements about the status of individual fish stocks are made on the basis of formal stock assessments reported in the RAM Legacy Stock Assessment Database (RAM) [@ricard2012]. The "assessed" fisheries in RAM represent roughly 50% of global capture as of 2020 [@hilborn2020], and represent our best estimates of the biological state of "assessed" fish populations around the globe.

That leaves roughly 50% of global landings, and the majority of global fisheries, as currently "unassessed". The SOFIA report includes unassessed stocks and bases its estimates for these fisheries on data-limited methods or qualitative expert opinion for each region where these stocks are distributed [@fao2020]. While these methods combined with local knowledge can provide good insight as to the status of unassessed fish stocks, the SOFIA assessment was designed in the 1970s based on the then available data and methods. With the surge in data availability, such as global assessments of management strength, trawl footprints, and fishing effort, the SOFIA global assessment methods for unassessed fish populations need to be updated to meet the demand for tracking progress of global fish populations towards the SDG goals.

We evaluate the ability of alternative assessment models to improve our estimate of the population status of unassessed fish stocks around the world. While these unassessed stocks are generally individually smaller and less economically valuable than the fish populations in the assessed category, collectively they are a vital source of food, employment, cultural value, and ecosystem services around the world. Numerous studies in recent years have put forward versions of "data-limited" models that have attempted to provide estimates for the global status of unassessed fish stocks lacking the data needed for formal stock assessment [@costello2012b; @costello2016a; @pauly2007; @rosenberg2018; @thorson2012a]. Due to data limitations, all of these global assessment efforts use forms of "catch-only" data-limited models (Free et al. 2020 [@free2020] and references therein). These models seek to infer stock status, for example biomass (*B*) relative to the biomass at maximum sustainable yield (*B~MSY~*), from characteristics of a fishery's catch history, and have been the foundation of many subsequent analyses of the future of global fisheries XX.

However, Free et al. 2020 [@free2020] demonstrated that these catch-only models can often produce both imprecise and biased estimates of current stock status in terms of *B/B~MSY~*. These issues become apparent when we consider some of the macro-level predictions made by these models. While the estimates *B/B~MSY~* and other reference points reported in RAM are themselves model outputs subject to their own non-trivial errors and biases, a simple benchmark is to compare the estimates of fishery status from RAM to those predicted by potentially less reliable catch-only models intended for use when insufficient data are available for a full stock assessment model.

Costello et al. 2016 [@costello2016a] found similar rankings of regions in terms of stock status as RAM, but their estimate of state of fisheries in the Mediterranean/Black Sea regions and Southeast Asia seem to be overly-optimistic, and the Northeast Pacific should be better by comparison. The superensemble method [@anderson2017a] used in Rosenberg et al. 2018 [@rosenberg2018] demonstrates the same problem, with stocks in Southeast Asia estimated as doing better than the Northeast Atlantic or Northeast Pacific. The Pauly 2007 [@pauly2007] catch-based approach finds the stocks of Southeast Asia in much better condition than those in the Northeast Pacific or Northeast Atlantic (Table.\@ref(tab:status-tab)). The non-RAM results in Table.\@ref(tab:status-tab) include both formally assessed and unassessed stocks, and as such we would expect them to differ broadly in their estimates of regional stock status, particularly since we might expect unassessed stocks to be less rigorously managed and by extension have poorer stock status. However, the lack of consistency across heavily assessed regions such as the Northeast Pacific, and the lack of contrast in stock status between heavily and lightly managed regions is concerning.

@@

tmp <- readr::read_csv(here("data","status-table.csv")) %>% 
  select(-`B/BMSY RAM`) %>%  
  left_join(o_by_a, by = c("FAO Area" = "fao_area")) %>% 
  left_join(state_space_results, by = c("FAO Area" = "fao_area")) %>% 
  arrange(`FAO % Overfished`) %>% 
  select(`FAO Area`,`FAO % Overfished`, everything())

 tmp %>% 
  knitr::kable(caption = "Estimates of B/B~MSY~ by FAO, RAM Legacy Stock Assessment Database through Hilborn et al. 2020, Costello et al. 2016, 
               Rosenberg et al. 2018, and Pauly 2007",
               booktabs = TRUE,
               format = "pipe") %>% 
  kableExtra::kable_styling(full_width = T, font_size = 9) %>%
  kableExtra::row_spec(0, bold = T)

@@

While each of these prior efforts at assessing global fisheries made important advances in our understanding of the global oceans, none have proven to be a consistently reliable means of estimating stock status of the worlds unassessed fisheries. In this paper we ask, can augmenting the types of catch-only models employed in global assessments to date with new broadly available data improve our understanding of the state of global fish populations? We use a flexible stock assessment software package, [sraplus](www.github.com/danovando/sraplus) to demonstrate how different data sources can be used to augment catch-only models at a global scale, and to evaluate how our perception of global stock status would vary depending on which sources of data we include. We show that improving our understanding of the world's unassessed fisheries depends on a redoubled effort at global data collection and synthesis.

@@

We compared estimates of stock status, defined by *B/B~MSY~* obtained from sraplus using various forms of broadly available data to the best-available estimates of *B/B~MSY~* available for those fish populations reported in the RAM database. All analysis were conducted in the R programming language [@rcoreteam2019]. Model fitting was conducted using Rcpp [@eddelbuettel2011] and Stan [@carpenter2017] implemented through Template Model Builder [@kristensen2016] by the tmbstan package [@monnahan2018]. The sraplus package is publicly available at [github.com/danovando/sraplus](https://github.com/DanOvando/sraplus), and all materials needed to fully reproduce this manuscript are available at [github.com/DanOvando/assessing-global-fisheries](https://github.com/DanOvando/assessing-global-fisheries). We encourage readers to explore the documentation available at the package website at www.github.com/danovando/sraplus. Here we describe the structure of the population model underpinning sraplus, the estimation models used, and the construction of priors used in this paper.

## Data Sources {.unnumbered}

We restricted ourselves to databases with wide coverage around the world, representing the types of information that can be leveraged in efforts to assess the status of global fisheries (Table.\@ref(tab:data)). While fishery-independent surveys are becoming increasingly available [@maureaud2020], they are not yet sufficiently distributed or accessible to serve as a foundation for global assessments of unassessed fisheries.

@@

## Experiment Structure {.unnumbered}

We assessed the ability of Fisheries Management Index (FMI), swept area ratio (SAR), and effort data to improve estimates of the state of global unassessed fish populations. We based this test on `r ram_comp_data$stockid %>% n_distinct()` stocks from RAM, covering `r n_distinct(ram_comp_data$isscaap_group)` broad taxonomic groups, with estimates of *B/B~MSY~* and greater than 25 years of continuous catch history. *B/B~MSY~* values from RAM are themselves estimates, not data, but they are the best available information on global stock status. We then paired the catch histories for these RAM stocks with regional-level SAR, FMI, and effort data. This process approximates a global-level assessment exercise, where data are available at regional levels, but not for specific fisheries.

We also estimated *B/B~MSY~* values of our candidate RAM stocks by using sraplus to fit to an abundance index drawn directly from RAM. We then fit a range of models utilizing different combinations FMI, SAR, and effort data, along with the CMSY method described in Froese et al. 2017 [@froese2017] (See Table.\@ref(tab:dat-desc)). We assessed performance using three metrics: median percent error (MPE, a measure of bias), median absolute percent error (MAPE, a measure of accuracy), and classification accuracy. Classification accuracy was calculated as the proportion of times that use of a given combination of data resulted in a stock being classified into the correct FAO status classification (one of underfished, maximally sustainably fished, and overfished).

@@

The core of sraplus is a Pella-Tomlinson [@pella1969] production model constructed in the manner of Winker *et al.* 2018 [@winker2018]. While models of these kinds abstract away many important details of fish biology and fleet behavior, they are the highest resolution model that the data types evaluated here will support. The purpose of sraplus is not to make substantial innovations in the fitting of surplus production models, but to provide a flexible tool for exploring the impacts of adding different kinds of data and priors on estimates of fishery status

@@

We allow for process error *p* (in the manner of the stochastic stock reduction analysis suggested by Walters *et al.* 2006 [@walters2006]). This allows the population dynamics to deviate from the exact values given by the Pella-Tomlinson operating model, while still conforming to the assumptions of this model on average. Incorporation of process errors is useful for two reasons: (1) when you have an abundance index, process errors can reduce bias arising from lack of fit in a deterministic SRA whenever dynamics are poorly explained by catch-history alone, and (2) with or without an abundance index (or other info), the stochastic portion is necessary to get good uncertainty intervals (i.e., with close to nominal coverage, see [@thorson2018]).

@@

In addition to the results from sraplus, we include a set of results produced by the default settings of the CMSY method [@froese2017]. For computational efficiency, we used a ported version of the CMSY model available at <https://github.com/DanOvando/portedcmsy>. The only modification made is to convert the underlying population model to C++ for faster computation. For each stock we used all the default options and priors provided and generated by CMSY, except for resilience, which was pulled from the vulnerability scores from FishBase accessed through `rfishbase` [@boettiger2012]. Vulnerability scores greater than 66 were scored as low resilience, between 33 and 66 medium resilience, and lower than 33 high resilience.

<!-- In the absence of an index of abundance CMSY and sraplus are conceptually the same: Both sample from prior distributions of life history and stock status conditional on the stock not collapsing given a catch history and a functional form for the population growth equation. The core different between CMSY and sraplus as utilized here then is that when we refer to CMSY we mean CMSY run using the default heuristics used by the algorithm to construct priors on initial, intermediate, and terminal stock status. In contrast, sraplus fit using for example FMI data uses the empirical models described here to generate priors on stock status. In addition, all sraplus runs use a prior-predictive tuning procedure to resolve issues related to Borel's paradox, described below. -->

@@

<!-- We can think of this as a binomial process in which we fit a model of the population, conditional on catches, to the fact that we know that population existed in each time step of the catch history. Beyond that though, baring additional information or model assumptions we have no way of knowing whether these catches represent a substantial proportion of a small population or a minuscule fraction of a massive population; all we know is that the population must always have been above zero. -->

Suppose that the only thing we observe from a fishery is a catch history. Assuming only Pella-Tomlinson population dynamics, the only thing we can learn from this catch history alone is the set of model parameters that ensure that the population still exists and never collapsed in the past. In the absence of any data to fit to, the a stock reduction algorithm  (SRA) works by assuming that we know current stock status, and then finds feasible parameters to satisfy that belief given a catch history and model structure. This creates a problem for the Bayesian nature of our analysis. Consider a production model with two parameters, a growth rate *r* and a carrying capacity *K*. Once we specify prior distributions on *r* and *K*, and then apply these distributions to our model (the shape of the production function along with the catch histories), we have implicitly provided a prior on the status of the stock in all time periods, since each unique combination of *r* and *K* together with the model and the catch history produces a deterministic stock status in each time step. Doing so places two priors on recent stock status: one implicit prior through the population parameter priors, and one explicit through the users perception of recent stock status, creating a problem termed Borel's Paradox (See Poole and Raftery 2000 [@poole2000] and references therein for a discussion of Borel's Paradox in a fisheries context).

<!-- This may seem like an academic concern, and indeed in our experience when the data are sufficiently informative the Bayesian version of our model subject to Borel's paradox produces effectively identical results to those produce by the same model fit by maximum likelihood.  -->

Borel's Paradox poses a particular problem  for the SRA version of sraplus. Due to the fact that there are more ways for a fishery to be relatively unexploited than for a fishery to be close to collapse but not collapsed, in this context Borel's Paradox causes the posterior distribution of stock status to be positively biased relative to the supplied prior (although other modeling choices can result in a net negative bias in stock status, @free2020). This process can also make it easy for users to accidentally supply very informative priors on stock status, without realizing that choices relating to population biology priors that may appear independent of stock status are in fact dictating the the posterior distributions of stock status resulting from the SRA algorithm.

We use an approximate solution to this problem here, similar in spirit to Bayesian melding [@poole2000]. Our solution amounts to a two-step sample-importance-resampling (SIR) algorithm. We first run the standard SIR algorithm as described above. We then break the resulting draws into bins based on terminal stock status, and calculate the mean sampling probability of each bin. The net result of this is that it allows users to place explicit prior on stock status, and then adjust their priors on life history parameters to reflect this prior, rather than creating a complicated and biased prior on stock status based on a mixture of explicit and implicit priors. See XX for a detailed explanation of this problem and our solution.

Along with allowing users to supply their own priors, the sraplus package contains three built-in methods for converting information on stock status from additional outside data into a form usable as a stock status prior by sraplus. We paired data on catch histories, swept area ratio (SAR), and fisheries management index (FMI) with estimates of stock status from the RAM legacy stock assessment database. We then trained a model of the general form $log(status) \sim normal(variable,\sigma)$ for each of these three data types. Given values of these variables for a new fishery, sraplus uses the fitted model to generate posterior predictive distributions of stock status based on these data, which can then be used as priors on stock status by sraplus for new fisheries. For example, given data on SAR or FMI scores, together with a catch history, sraplus uses these regressions to convert those SAR and FMI values into priors on B/B~MSY~ or F/F~MSY~ in the most recent year of the fishery usable by sraplus. See XX for a detailed explanation of this process.

All prior regression models where tested by out-of-sample predictive power, and where competing models were considered the final model was chosen by leave-on-out validation [@vehtari2017]. The final models are intended as a reasonably robust means of translating available data (catch histories, FMI, and SAR values) into a form usable by sraplus. Given the scope of this analysis, we do not claim that the regressions used here are the best possible model relating these data with the fishery status indicators of interest. Rather, each regression was tested to ensure that it is unlikely that, given the same data, an alternative model would perform substantially better than those presented here.

We performed a value-of-information (VOI) assessment to determine what types of data may be most beneficial to acquire at a global scale if we are to improve our knowledge of the state of global fisheries. The VOI analysis was performed by using sraplus to generate estimates of *B/B~MSY~* for stocks in RAM, and comparing the estimated values to the values reported in RAM. We generate fits for 3000 combinations of a RAM stock and available data. For any one draw, we randomly sample a RAM stock and a list of available data and data quality. For example, we might sample stock *A* with information on recent fishing mortality rates for the first iteration, and stock *A* again for the second iteration but now with information on recent fishing mortality rates and a recent index of abundance. The result is a set of model performance estimates where the characteristics of the stock and the data made available to the model are randomized.

We first present a case study demonstrating how different kinds of data can lead to different conclusions about stock status. From there, we assess the performance of models fit using different kinds of broadly available data: combinations of catches, effort, Fisheries Management Index scores, and swept area ratio (SAR) values (Table.\@ref(tab:data)).

As a case study, we selected `r n_distinct(exs$stockid)` stocks for which we have stock specific FMI and SAR scores. We then paired effort data at the resolution of year, country, and FAO statistical area from Rousseau et al. 2019 [@rousseau2019] to each stock. As a benchmark, we first estimated stock status for these case study fisheries using the CMSY [@froese2017] method, as this has become one of the most widely used catch-only models currently available. We then used stock-specific data on SAR and FMI to generate priors on F/F~MSY~ for each of the stocks, which were then passed to sraplus. Lastly, we used the reconstructed effort data [@rousseau2019] to create an index of abundance for each stock, and estimated stock status by fitting to this index while using priors on fishing mortality rates informed by each stock's FMI and SAR values. While CMSY systemically overestimated fishing mortality rates and underestimated stock status, use of the SAR, FMI, and effort data produced substantially more accurate results for both B/B~MSY~ and F/F~MSY~ (Fig.\@ref(fig:cs-plot)).

We next assessed the ability of FMI, SAR, and effort data to improve estimates of global stock status. Overall the sraplus estimates of B/B~MSY~ resulting from using the RAM data were relatively accurate and unbiased at a macro level (MPE `r percent(perf$mpe[perf$"Data Used" == "ram-data"])`,MAPE `r percent(perf$mape[perf$"Data Used" == "ram-data"])`, accuracy = `r percent(perf$accuracy[perf$"Data Used" == "ram-data"])`, Table.\@ref(tab:perf-tab), Fig.\@ref(fig:mpe-map)-\@ref(fig:acc-map)). This exercise tells us that given sufficiently high quality data, a surplus production model such as sraplus is reasonably capable of reproducing the global state of fisheries as understood from formally assessed fisheries.

Performance limitations in assessments of global fishery status then are likely to arise less from model misspecification than from the quality of the data themselves. This becomes clearer once we consider the performance of sraplus models fit to combinations of broadly available datasets. Many of the datasets used produced similar levels of bias as the RAM data (Table.\@ref(tab:perf-tab)). However, this is somewhat an artifact of the data. The status of most stocks in RAM is also relatively good, with recent B/B~MSY~ values generally near one. This means that a model that more or less reproduces the global average of stock status will be relatively unbiased on average, but imprecise. Focusing on MAPE instead, the error of the models jumps dramatically as soon as data other than RAM are used, to a minimum value of `r percent(perf$mape[2])` and a maximum of `r percent(max(perf$mape))`. The mean accuracy of the sraplus models across all non-RAM data fits was only `r percent(mean(perf$accuracy[!perf$"Data Used" %in% c("ram-data","guess")]))`. Note that there are only three bins in the FAO stock status classifications, and a "model" that randomly assigns a stock to a status category would have a mean accuracy of `r percent(perf$accuracy[perf$"Data Used" == "guess"])`.

The poor global performance of models fit to currently accessible broadly available datasets, relative to models fit to data from RAM, indicates that we must prioritize collection and aggregation of new or underutilized data sources if we are to improve our estimates of global fishery status. What sources of data might provide the greatest value in improving our estimates of global stock status? We used sraplus together with the RAM database to estimate the value of information, measured as the average reduction in root mean squared error, resulting from use of different kinds of data (Fig.\@ref(fig:voi-plot)). Having access to estimates of *F/F~MSY~* reduced model error in proportion to the number of years for which *F/F~MSY~* values are available. Interestingly though, having access to only an accurate estimate of F/F~MSY~ in the most recent year was extremely informative, reducing error on average by 15%, on par with an estimate of recent *B/B~MSY~* itself. While having access to complete index of abundance, such as a fishery independent survey, was on average able to reduce error relative to a baseline catch-only heuristic, using only the most recent quarter of the available abundance index actually increased error on average. We may have to wait many years for new surveys to provide substantial improvements in status estimates, or work to expand access to long-running existing surveys that have yet to be fully utilized in fisheries assessment [@maureaud2020].

Global-level assessments of fish populations are critical for guiding management agendas for the world's oceans, and tracking indicators such as the United Nations Sustainable Development Goals. Despite this need, and despite advances in stock assessment methods and available data, we show that our understanding of the world's fish stocks remains uncertain in many parts of the world. While in some cases addition of globally available data, such as quality of fisheries management or effort reconstructions, provided value above and beyond catch histories alone (Fig.\@ref(fig:cs-plot)), at the global level models fit using each of the available datasets, besides the RAM-derived indices, generally produced biased and imprecise estimates of fish stock status (Table.\@ref(tab:perf-tab)).

We chose to test the performance of methods against estimated values reported in RAM. A reasonable critique of this choice is that unassessed stocks, on which these methods would actually be used, are likely to have vastly different dynamics than the heavily managed fish populations represented in RAM. The clear and systemically poor performance of the methods tested here when applied to RAM stocks would require though that these methods have massively lower bias and higher accuracy for unassessed fisheries than RAM stocks if they are to be expected to provide reasonable picture of global fish populations. Free et al. 2020's [@free2020] simulation testing of the types of methods included here suggests that this is unlikely to be true.

Our results do not imply that the kinds of broadly available data presented here are not valuable under the right conditions. The FMI and SAR based priors are an improvement over catch-only models in applicable situations (i.e. those that sufficiently resemble the data on which the regressions were trained, Fig.\@ref(fig:cs-plot)). Effort data such as those reconstructed by Rousseau et al. 2019 [@rousseau2019] can help distinguish between regions with similar catch histories but different large-scale effort trajectories, and may be quite useful as indices of abundance for areas with good knowledge of rates of evolution of fishing technology and a broadly selective fishing fleet. Despite not adding a great deal in terms of performance at the global scale, swept-area-ratio was the strongest predictor of F/F~MSY~ of any of the datasets we explore on an individual stock basis, with a Bayesian R^2^ value of `r round(mean(rstanarm::bayes_R2(sraplus::sar_models$fit[[3]])),2)`.

We must also prioritize collection and curation of fish population survey data worldwide. Repositories of fishery-independent survey data would be immensely beneficial, such as those maintained by [FishStat](https://james-thorson.github.io//). Recent research confirms that there are bottom trawl data to support analysis of biomass-trends since 2001 and potentially earlier in many regions [@maureaud2020], and survey data are available for more stocks than have stock assessments. Effort reconstructions such as those utilized here may help create fishery-dependent abundance indices in some instances, and going forward datasets such as those compiled by [Global Fishing Watch](https://globalfishingwatch.org/) in combination with the reconstruction approaches of [@rousseau2019] might allow us to construct and use timeseries of fishing effort specific to particular areas, fleets, and species complexes

Expanded training of fisheries scientists around the globe is another critical need. Even were we to dramatically expand the amount and types of data available for global assessment, individual fisheries and regions will need to make informed decisions about which sources of data may be applicable and which not, and to critically evaluate the results of any model based on local expertise. This is why stock assessments even in data-rich fisheries are not an automated process; the real challenge is often not in fitting a model to data but in understanding how best to use the data and the quality and limitations of the model used. Empowering a global network of fisheries scientists through training and peer-support would help local experts make the most of available data, ensure the reliability of newly collected data, and improve the interpretation of assessment results.

The coming decades are a critical time for the future of fisheries and ocean health. Achieving the United Nations Sustainable Development Goal 14 for the conservation and sustainable use of the world's oceans depends on our ability to effectively assess the status of fish stocks around the world. The RAM Legacy Stock Assessment Database combined with the FAO's expert elicitation of status for select stocks have dramatically improved our understanding of global fisheries in recent years. However, this process still leaves a substantial number of fisheries and proportion of global catch unassessed. Numerous catch-based data-limited approaches have attempted to fill that gap, and while these efforts have advanced our knowledge and interest in unassessed fisheries, none have yet been able to provide a clear solution to this problem.

<!-- Our results show that improving estimates of global stock status depends on investing in an improved and expanded global network of fishery data and fisheries scientists. -->

The demonstrated lack of strong information on stock status within catch histories alone means that differences in catch-only model structure and core assumptions between different assessment efforts can produce starkly contrasting conclusions on global stock status, leading to debates that are inconclusive as they are inherently driven by assumptions. The need is to develop better datasets and approaches on the ground that can lead to more conclusive inferences about the state of global unassessed fisheries. The FAO is leading efforts to increase technical capacity and monitoring and evaluation infrastructure to better inform fisheries management in places with limited data. Such projects stand to provide a better picture of fishery status at global and local scales, furthering our ability to meet the UN SDG targets. Our results emphasize the urgency and rationale for building the infrastructure and capacity that can lead to better marine resource management globally . Catch-only methods may be a good starting point when applied properly, but achieving meaningful improvements in the assessment and management of global unassessed fisheries will depend on expanded collection of targeted data types, active management, and local capacity building.

## Prior Predictive Tuning {.unnumbered}

@@
